\chapter{Method Implementation And Results}

In this chapter the tools used, the software infrastructure and method implementation and the results collected during the analysis are reviewed.

A DDPG Agent\footnote{An agent whose actions are taken by a neural network trained with the reinforcement learning algorithm DDPG} was trained to drive in an urban environment. Checkpoints of the network's state during the training were recorded for the purpose of comparison. These checkpoints of the network were then tested with and without a simple safety-monitor in order to provide a new point of view to study AV's behaviours.

\section{Tools and software}

\subsection{Carla Simulator}

In order to have a realistic environment, with accurate physics simulation and data sensors, the open-source simulator CARLA\cite{carla}, developed by researchers at the University of Barcelona, was used. This simulator was developed with the purpose of offering an environment where AI agents can be trained to drive, with high control of the simulation parameters and the simulation of realistic sensor, which can be tuned to increase or decrease data quality, or to inject faults.\newline
CARLA is developed with a client-server architecture in mind. The \textsl{server} is basically a game, developed with \textsl{Unreal Engine 4} in C++. C++ performances are with no doubt essential to the functionality of the server: not only the environment must be simulated (including movements of pedestrians/vehicles, weather simulation\dots), but also all the data needed from the sensors attached to the system.\newline


======================================================================================

IMMAGINE CARLA

======================================================================================

CARLA is currently at version 0.9.7 and huge improvements are done at every release, gaining more attention from the experts for its realism. Unfortunately, when this study started, CARLA 0.9 was recently released and the tools needed for our work couldn't be found online. Thanks to the quantity of work done for the last \textit{stable} version of CARLA, 0.8.4 was used at first.\newline
Versions prior to 0.9 have some limitations on the control one has of the simulations parameters and on the data collectable from it. This doesn't impede our study, but of course limited in some way the informations on the environment and system. Some of these problems are still present in later versions of the simulator, but most of them were solved in the transition from 0.8 to 0.9.\newline\newline
One of the main problems found was with the coordinate systems. Before version 0.9, developers were using UE4's default coordinates system which is left-handed, while the standard is considered to be right-handed. This looks like not a big deal since things could be easily solved by applying a transformation matrix. However, due to performance issues (a Python client should do the real-time processing of \textsl{loads} of data at each timestep, resulting in considerable slowdowns as a result of all the processes running at the same time), it was decided to stick with the developers' decision and convert the data during analysis phase.

Unfortunately, this version of CARLA has only 4 sensors available, which were all used during the experiments. They can be easily accessed via the Python APIs provided:

\begin{itemize}
	\item Cameras
	\begin{itemize}
		\item The \textsl{scene final} camera provides a view of the scene (just like a regular camera)
		\item The \textsl{depth map} camera assigns RGB values to objects to perceive \textsl{depth} of the environment
		\item A \textsl{semantic segmentation} is used to classify different objects in the view by displaying them in different colors, according to the object's class
	\end{itemize}
	\item Ray-cast based Lidar
	\begin{itemize}
		\item Light Detection and Ranging is use to sense the environment and measures distance from objects by illunating the target with laser beams and measuring the time reflected light needs to "go back" to the sensor
	\end{itemize}
\end{itemize}

The three cameras were used during the training phase of the network. Three \textsl{scene final} cameras are attached to the car to actually \textsl{see} the environment (one on the front and one per side). The \textsl{depth map} camera allows the car to get a colormap of the distances from objects in the scenario.
The \textsl{semantic segmentation} provides image classification features by querying the server for ground-truth values. This is with no doubt a semplification of a real system, where the most powerful image-classification softwares are essentially other neural networks trained separately. At the same time a misclassification can be considered as an error of the control system: if the safety monitor detects the possible hazard it will not "correct" the misclassification but it must react fast and safely to avoid the possible consequences of it, therefore this simplification won't have an impact on the overall method.\newline

A ray-cast based Lidar is the only other sensor available for this version of CARLA. Parameters of this sensor can be easily tuned to simulate real lidars such as the \textsl{Velodyne LiDAR} or to simulate faults such as low data quality, noisy data or data loss\dots This data are generated using the \textsl{Point Cloud}\cite{pointcloud} format

In the simulations, due to the high hardware resources requirements to simulate a real LiDAR, a slightly modified version of the \textsl{Velodyne64 LiDAR} is implemented with the following parameters:

\begin{itemize}
	\item Channels = 64
	\begin{itemize}
		\item The number of laser beams used by the system. These lasers are distributed over the vertical axis. The more the lasers are, the more accurate will be the scannings
	\end{itemize}
	\item Range = 75m
	\begin{itemize}
		\item Lasers' range in meters
	\end{itemize}
	\item Rotation Frequency = 15 Hz
	\begin{itemize}
		\item This parameters define the rotation frequency (in Hz) of the scanning beams.
	\end{itemize}
	\item Points Per Second = 1.000.000
	\begin{itemize}
		\item The actual number of points generated each frame by the sensor
	\end{itemize}
	\item Vertical FOV bounds (height = 24m, low = -2m. Distances are relative to the position of the sensor)
	\begin{itemize}
		\item Maximum and minimum height of the scannings
	\end{itemize}
\end{itemize}

The simulator provides Python APIs not only to modify sensors, but also to have a great control on what is being simulated, such as seeds definition for the spawning points and the behaviours of pedestrians and vehicles, and on the state of \textsl{"actors"} in the scene such as their position, their speed\dots. All these data are directly provided by the simulator with ground-truth values. These kind of measurements can be simulation-related, such as the simulation time-step, or the FPS. Actors-related measurements include for example vehicles' speed, intensity of collisions (if any) and the 3D acceleration vector.\newline

While this tool was being tested a problem was found with the LiDAR sensor data that is still not resolved.\cite{lidarbug} 
The bug resulted in the vehicles' bounding boxes to be distorted when moving, providing very poor data.
After some research an improved version of the simulator was found, where the bounding boxes for each vehicle were redefined to provide accurate data.\cite{carlapro} As of today, developers are working on this issue but it is still unresolved without manual intervention on the source code.


\subsection{Controller Implementation}

The main component needed for the Controller implementation is the neural network that will be in charge of deciding the action to perform to drive the car. We looked for a framework with the following characteristics

\begin{itemize}
	\item[1] Training code must be available
	\item[2] No critical issues in the codebase
	\item[3] Provide an environment to interface the network with CARLA
	\item[4] Provide a default training strategy
\end{itemize}

After analyzing all the machine-learning related projects for CARLA, we chose the reinforcement-learning framework \textsl{Coach}, developed by \textsl{Intel AI Lab}.\cite{coach}

This framework satisfies all the requirements listed above: it is distributed as an \textsl{editable Python package}. The development team behind this project assure us the quality of the product. Moreover, several presets are available as a starting point. Two of these presets offer an interface to the CARLA simulator, as well as a default training strategy, which is exactly what we required.

These presets implement the \textsl{Deep Deterministic Policy Gradient} algorithm, proposed in 2015.\cite{ddpg} This algorithm proved to perform well in tasks such as car driving, as shown in the original paper, and it's specifically adapted to perform in environments with continuos action space, such as the one we are considering.

\begin{itemize}
	\item[P1)] The first preset uses a single, front camera to perceive the environment, and the other data augmentation cameras provided by CARLA. The agent resulting from using this preset will likely not have any sense of depth created by the regular camera and will rely solely on the depth camera
	\item[P2)] This preset uses all the cameras available in CARLA and have a much more interesting architecture since the car is equipped with three regular cameras (Front, Left, Right) and all the data augmentation cameras available in CARLA, listed in the previous section
\end{itemize}

The usage of the data augmentation cameras (depth and semantic segmentation) allows us to ignore objects misclassifications (e.g. labeling a pedestrian as a light pole) as well as some other easinesses such as computing distances from objects. As pointed in the previous section, this is indeed a simplification of a real architecture, where object detection modules are neural networks themselves, that must be trained in advance and tested separately. However, again, a misclassification will probably cause the Controller to behave in unexpected manners that the Safety Monitor must be able to detect and possibly correct.

Unfortunately, the first preset suffers from an issue that causes the car to stop throttling, while being positively rewarded\cite{coachIssue}. It would have been interesting to test this preset as well, to understand what's the impact on the effectiveness of the Safety Monitor with respect to the quantity of data available to the network.\newline

It's important to notice that our goal \textsl{is not} to build the \textsl{"perfect"} agent, nor the perfect autonomous cars. The codebase was studied but, due to reasons of time, we could not inspect all the details of the provided implementation. This framework was used as an example to demonstrate the concepts described in this work.


\subsection{Safety-Monitor Implementation}

In order to start the experimental activity, we needed a software capable of processing LiDAR data to map the environment and sense the obstacle. Moreover, this software needs to be \textsl{real-time} and capable of interacting with CARLA. The latter is obvious: whenever an alert is raised, the \textsl{"brake"} command must be sent to the CARLA server. The first required some reasoning: one could think to record the LiDAR data in advance and then run the simulations feeding the Monitor with these data. Unfortunately this approach can not work: this comes from the fact that we are not interested in having a 100\% prediction accuracy, but also in the \textsl{goodness} of the safety-measure adopted by the Monitor. The same measurement in two different simulations could give \textsl{slightly} different results, that could cause the Monitor to behave in a completely different way than the way it could have behaved using the \textsl{real time} data produced during the simulation. If the accuracy of measurements is strictly dependent on the simulator used for the experiments (CARLA is improving at every release on this side), we could not ignore the effects that a brake can have on a single run.

In order to develop a decent Safety-Monitor, a research on the best, \textsl{"not-neural-network"} techniques was conducted, as well as a research on the open-source instruments available. The choice fell on the \textsl{Point Cloud Library}\cite{pcl}, an open-source library specifically dedicated to the processing of Point Cloud data, developed using C++ to achieve great performances when processing huge amount of data.
This library was first released in 2011\cite{pclwiki} and it's been improved at every release, also thanks to the big community testing and debugging the new features.

This library offers a multitude of functions that implements the most known techniques for Point Cloud processing. The steps to follow in order to develop an object-detection module are the following:

\begin{itemize}
	\item[1)] Downsampling
	\begin{itemize}
		\item[$\rightarrow$] The data generated by a single scan can contain more hundreds of thousands records, with possibly a lot of redundancy and noise. A downsample is usually necessary as a preliminary step to discard all the \textsl{"useless"} data. After some reasoning, the Voxel Grid Filter was chosen for this step
	\end{itemize}
\end{itemize}

\begin{itemize}
	\item[2)] Ground Segmentation
	\begin{itemize}
		\item[$\rightarrow$] After the downsampling (if needed), the first mandatory step is to filter the data that are useless for the purpose of object detection, such as points relative to the ground. These point must be filtered in order to separate the ground from the objects we want to detect.
		In our implementation this is done using the RANSAC\footnote{Random Sample Consensus} algorithm\cite{ransac}, a technique to separate \textsl{"inliers"} (i.e. data whose distribution is characterized by the model's parameters) from \textsl{"outliers"} (i.e. data that don't have a representation in the chosen model)
	\end{itemize}
\end{itemize}

==================================================================================

IMMAGINE PRE-POST SEGMENTATION

==================================================================================

\begin{itemize}
	\item[3)] Clustering
	\begin{itemize}
		\item[$\rightarrow$] This final step is required to effectively identify what is an object in the scene, and what data points are relative to the same object. This can be done using clustering algorithms based on the distance among points. We chose the Euclidean Clustering Algorithm, a range search algorithm based on the euclidean distance between points, with the assumption that \textsl{very dense} points represent the same object
	\end{itemize}
\end{itemize}


==================================================================================

IMMAGINE CLUSTERING

==================================================================================

\begin{itemize}
	\item[4)] Object Tracking and Avoidance
	\begin{itemize}
		\item[$\rightarrow$] Objects detected in the first three steps must be tracked over time in order to recognize if two objects observed in two consecutive steps, are actually the \textsl{same} object, usually using physical models to predict their behaviour over time, such as the Kalman Filter, and combine this model with a failure-prevention routine
	\end{itemize}
\end{itemize}

Unfortunately, implementing a Kalman Filter for such a complex model would have been too much time-consuming, alting our study. Therefore we simplified the object-detection and the failure-prevention routine in this way:

\begin{itemize}
	\item Only data \textsl{in front of} the car are recorded. This is a huge semplification of the model, as the Monitor now can only detect obstacles ahead of the car. However, this will for sure have an impact on the effectiveness of the Safety Monitor, but the concepts explained in the previous section still holds
	\item The failure-prevention routine implemented takes inspiration from the Responisbility-Sensitive Safety model proposed by Mobileye, an Intel company.\cite{rss}
	If an object is detected, its speed relative to the system is computed. If the distance travelled by the system in 1 second  plus the braking distance is greater than the distance travelled by the object plus the distance between the system and the object, a safety brake is applied.
\end{itemize}

Since the Safety-Monitor is developed in C++, both for taking advantage of the Point Cloud Library and due to performance reasons, an infrastructure to exchange data with CARLA was developed.

The software is composed of a C++ server that receives LiDAR Point Clouds from the CARLA client. This data are processed as in the steps above (with the semplification described) and, if needed, an alert is sent back to the client. If the message received by the Monitor is an alert, the actions of the Controller are ignored and a brake is performed.

The object detection module is inspired by an open-source project developed by Engin Bozkurt\cite{LOD}. The codebase was deeply modified to adapt the detection algorithm parameters to fit our needs and to interact with CARLA.


\section{Experimental Activity}


LEGGERE GROUND-TRUTH PERMETTE DI IGNORARE ERRORI DI MISCLASSIFICATION DEGLI OGGETTI

DISCUTERE LE STRATEGIE DI TRAINING

\section{Results}

METTI QUI I NUMERINI CHE TI SONO VENUTI